---
title: "lazyeval exercises"
author: "Tristan Mahr"
date: "June 18, 2016"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, comment = "#>")
library("lazyeval")
```

## Part 1

> Create a wrapper around lm() that allows the user to supply the response and predictors as two separate formulas.

```{r}
lm2 <- function(response, predictor, df) {
  # I can't figure out how to uq both at once within a formula that uses a
  # formula (~ lm(y ~ x)), so I'll plug them into a formula one by one.
  f <- ~ y ~ x
  f_lhs(f) <- uq(response)
  f_rhs(f) <- uq(predictor)

  f_eval(~ lm(uqf(f), df))
}

lm2(~ mpg, ~ cyl, mtcars)

lm2(~ mpg, ~ hp, mtcars)
```

I don't know how I would incorporate `...` yet. I'm not that far in the vignette.

***

> Compare and contrast f_eval() with with().

Both can use names outside of the data.

```{r}
cyl_sq <- mtcars$cyl ^ 2

with(mtcars, coef(lm(mpg ~ cyl + cyl_sq)))

f_eval(~ coef(lm(mpg ~ cyl + cyl_sq)), mtcars)
```


I thought `with` would work better with blocks of multiple lines of code, so 
`f_eval` can do those too.

```{r}
with(mtcars, {
  log_wt <- log(wt)
  mean(log_wt)
})

f_eval(~ {
  log_wt <- log(wt)
  mean(log_wt)
}, mtcars)
```

The main difference is that `f_eval` provides an unquoting/interpolation step in
the middle.

```{r}
some_var <- ~ wt
with(mtcars, mean(uq(some_var)))

f_eval(~ mean(uq(some_var)), mtcars)
```

`with` jumps right into evaluation (`eval`).

```{r}
with.default
```

`f_eval` modifies the expression with `f_interp` before calling its evaluation
step (`eval_expr`).

```{r}
f_eval
```


*** 

> Why does this code work even though f is defined in two places? (And one of them is not a function).
> 
> ```{r}
> f <- function(x) x + 1
> f_eval(~ f(10), list(f = "a"))
> #> [1] 11
> ```

The same reason this can work coherently:

```{r}
log(100)
log <- 100
log(100)
log(log)
```

I don't know the low-level eval/apply steps in detail, but the interpreter looks
for a function with the name `f` when evaluating a function call by `f`, so it
skips over the piece of data named `f` that is not a function.


```{r}
f <- function(x) x + 1
f_eval(~ f(10))
f_eval(~ f(10), data = list(f = "a"))

# Put a function in the data
f_eval(~ f(10), data = list(f = sqrt))
```

